---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
```
# traipse


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/Trackage/traipse.svg?branch=master)](https://travis-ci.org/Trackage/traipse)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/7v59mpmj4lqw2b0q/branch/master?svg=true)](https://ci.appveyor.com/project/mdsumner/traipse)
[![Codecov test coverage](https://codecov.io/gh/Trackage/traipse/branch/master/graph/badge.svg)](https://codecov.io/gh/Trackage/traipse?branch=master)
<!-- badges: end -->

The goal of traipse is to provide shared tools for tracking data, for common metrics of distance, direction, and speed. The package includes the following functions which are always assumed to operate on input locations in longitude latitude, and input date-times in R's `POSIXt` class. 

* `track_distance()` for distance in metres
* `track_angle()` for internal angle in degrees
* `track_turn()` for relative turn angle
* `track_bearing()` for absolute bearing
* `track_time()` for duration in seconds
* `track_speed()` for speed in metres per second
* `track_distance_to()` for distance to location
* `track_bearing_to()` for bearing to location

Distances are always returned in metres, directions and angles are always returned in degrees. Absolute bearing is relative to North (0), and proceeds clockwise positive and anti-clockwise negative `N = 0, E = 90, S = +/-180, W = -90`. 

Time is always returned in seconds, and speed in metres per second. 

There is a huge fragmentation of tools and approaches in R for tracking, see the Spatio-Temporal 
Task View section for examples. 

We would like to have a simple core package to provide the most commonly used metrics. We assume `geodist` and `geosphere` as good examples of core packages for the underlying tool. These both apply the modern 

## Installation

Not available on CRAN yet. 

You can install the  development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Trackage/traipse")
```
## Example

This is a basic example which shows you how to calculate ellipsoidal distance and turning angle for a data set of tracks. 

First, calculate without any groupings - we definitely don't want this as there are three separate tracks within our data set identified by `id`. (No ordering is a applied other than the order the rows occur).

```{r example}
library(traipse)

library(tidyverse)
## there's no grouping here - we haven't gotten our data organized yet
trips0 %>% mutate(distance = track_distance(x, y), angle = track_angle(x, y))

```
Now run the same metrics but do it with respect to the grouping variable `id`. 


```{r example-group_by}
## now we group by id (also need arrange by date to be sure ...)
metric <- trips0 %>% group_by(id) %>% mutate(distance = track_distance(x, y), 
                                             angle = track_angle(x, y),
                                             turn = track_turn(x, y), 
                                             bearing = track_bearing(x, y), 
                                             duration = track_time(date),
                                             speed = track_speed(x, y, date), 
                                             distance_to = track_distance_to(x, y, 147, -42), 
                                             bearing_to = track_bearing_to(x, y, 100, -42)) 
metric 

metric %>% 
  ggplot(aes(x, y, cex= 1/angle)) + 
  geom_point() + 
  geom_path(col = rgb(0.2, 0.2, 0.2, 0.2))

metric %>% 
  ggplot(aes(x, y, colour = distance_to)) + 
  geom_point() + geom_label(data = data.frame(x = 147, y = -42, distance_to = 0), 
                            label = "home")

metric %>% 
  ggplot(aes(x, y, colour = bearing_to)) + 
  geom_point() + geom_label(data = data.frame(x = 100, y = -42, bearing_to = 0), 
                            label = "home")


```

Using the bearing and distance now reproduce the track as *destination point* segments. 

```{r dest-point}
plot(metric[1:10, c("x", "y")], type = "b", lwd = 10, col = "grey")
dest <- geosphere::destPoint(metric[1:10, c("x", "y")], 
                             b = metric$bearing[1:10], 
                            d = metric$distance[2:11])
arrows(metric$x[1:10], metric$y[1:10], dest[1:10,1], dest[1:10,2], col = "firebrick", lwd = 2)
```



## Data are assumed to be sensibly organized

Note above that we provided a *grouping ID* for when we have separate trips within the same data set. There's nothing to stop from calculating distances when the arrangement of records does not make sense, but this is your responsibility. If missing values are present, or times are out of order, or include zero-length time durations, or movement backward in time there are aren't any checks. 

The idea is for developers to be able to use these tools however they like. 

---

Please note that the 'traipse' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

